Enforced Invariant Confluence

### 2016-11-24 doc/counters.tex
Added some more todo proofs.
Tidying up diagrams.
Tidied up more.
Added proofs relating o-isafety and i-strength*

A \emph{database state} is a total function $D: \var \to \ints$.  We say a
database state $D$ \emph{satisfies invariant} $I$, denoted $I(D)$, if $I$
evaluates to true after all variables in $I$ have been replaced by their
mapping in $D$.

@@

\subsection{\iconfluence{} Alternatives}
\newcommand{\ipreservation}{$\iinvariant$-preservation}
\newcommand{\ipreserving}{$\iinvariant$-preserving}
\newcommand{\isafety}{$\iinvariant$-safety}
\newcommand{\isafe}{$\iinvariant$-safe}

\begin{itemize}
  \item \todo{Clarify these rambling thoughts.}
  \item \todo{Add $k$-\isafety to the figure.}
  \item \todo{Find something between \isafety{} and \iconfluence.}
  \item \todo{Add two state-based proof rules from strong enough paper limited
              to causal consistency.}
  \item \todo{Prove state-based proof rules imply iconfluence.}
  \item \todo{Figure out relationship of proof rules to this stuff.}
  \item \todo{1-iconfluence with derived txns to start}
\end{itemize}

@@

  \item \textbf{k-\isafety.}
    $\forall D, D', B, C, c.\>
       C = \denote{B}(D) \land
       |B| \geq k \land
       I(B @ D') \implies
       I(D \circ C \circ \denote{c}(D' \circ C))$

@@

  \item \textbf{1-\iconfluence.}
    $\forall D, c_1, c_2.\>
       I(D) \land
       I(D \circ \denote{c_1}(D)) \land
       I(D \circ \denote{c_2}(D)) \implies
       I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$
\end{itemize}

\begin{claim}
  $\text{\iconfluence} \implies \text{$k$-\iconfluence}$
\end{claim}
\begin{proof}
  \todo{Trivial.}
\end{proof}
\begin{claim}
  $\text{1-\iconfluence} \centernot\implies \text{$k$-\iconfluence}$
\end{claim}
\begin{proof}
  See \clmref{wimp-1-iconfluence}.
\end{proof}

\begin{claim}
  $k\text{-\iconfluence} \implies \text{1-\iconfluence}$
\end{claim}
\begin{proof}
  \todo{Trivial.}
\end{proof}
\begin{claim}
  $\text{$k$-\iconfluence} \centernot\implies \text{\iconfluence}$
\end{claim}
\begin{proof}
  See \clmref{wimp-1-iconfluence}.
\end{proof}

\begin{claim}
  $\text{\isafety} \centernot\implies \text{\ipreservation}$
\end{claim}
\begin{proof}
  \todo{Trivial.}
\end{proof}
\begin{claim}
  $\text{\ipreservation} \centernot\implies \text{\isafety}$
\end{claim}
\begin{proof}
  \todo{Not sure.}
\end{proof}

@@

  \begin{tikzpicture}
    \node[] (0-isafety) at (0, 0) {0-\isafety};
    \node[] (ipreservation) at ($(0-isafety) + (30:3)$) {\ipreservation};
    \node[] (iconfluence) at ($(0-isafety) + (-30:3)$) {\iconfluence};
    \node[] (k-iconfluence) at ($(iconfluence) + (4, 0)$) {$k$-\iconfluence};
    \node[] (1-iconfluence) at ($(k-iconfluence) + (4, 0)$) {1-\iconfluence};

    \draw[impl] (0-isafety) -> (ipreservation);
    \draw[impl] (0-isafety) -> (iconfluence);
    \draw[impl] (iconfluence) -> (k-iconfluence);
    \draw[impl] (k-iconfluence) -> (1-iconfluence);
  \end{tikzpicture}

@@

  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{d1}{0, 0}
      \inode[label=above:$D'$]{d2}{1, 0}
      \nnode{b1}{0, -1}
      \nnode{b2}{1, -1}
      \draw[txn] (d2) -- (b2) node[midway, right]{$t(D')$};
      \draw[dtxn] (d1) -- (b1) node[midway, left]{$t(D')$};
    \end{tikzpicture}
    \caption{1-\isafety}
    \label{fig:1-isafety}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{a}{0, 0}
      \nnode{b}{0, -1}
      \draw[txn] (a) -- (b) node[midway, left]{$t(D)$};
    \end{tikzpicture}
    \caption{1-\ipreservation}
    \label{fig:1-ipreservation}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{a}{0, 0}
      \inode[label=left:$D_{t}$]{l1}{$(a) + (240:1)$}
      \inode[label=right:$D_{u}$]{r1}{$(a) + (300:1)$}
      \nnode{b}{$(l1) + (300:1)$}
      \draw[txn] (a) -- (l1) node[midway, left] {$t(D)$};
      \draw[txn] (a) -- (r1) node[midway, right] {$u(D)$};
      \draw[dtxn] (l1) -- (b) node[midway, left] {$u(D)$};
      \draw[dtxn] (r1) -- (b) node[midway, right] {$t(D)$};
    \end{tikzpicture}
    \caption{1-\iconfluence}
    \label{fig:1-iconfluence}
  \end{subfigure}

@@

    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D_0$]{l1}{0, 4}
      \nnode{l4}{0, 0}

      \inode[label=above:$D_0'$]{r1}{1, 4}
      \inode[label=right:$D_1'$]{r2}{1, 3}
      \inode[label=right:$D_{n-1}'$]{r3}{1, 1}
      \nnode{r4}{1, 0}

      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
      \draw[txn] (r2) -- (r3) node[midway, right]{
        $\begin{array}{l}
          \denote{c_2}(D_1'), \\
          \ldots, \\
          \denote{c_{n-1}}(D_{n-2}'), \\
        \end{array}$
      };
      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_n}(D_{n-1}')$};

      \draw[dtxn] (l1) -- (l4) node[midway, left]{
        $\begin{array}{l}
          \denote{c_1}(D_0') \circ {} \\
          \denote{c_2}(D_1') \circ {} \\
          \cdots \circ {} \\
          \denote{c_n}(D_{n-1}')
        \end{array}$
      };
    \end{tikzpicture}

    \caption{\isafety}
    \label{fig:isafety}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering

    \begin{tikzpicture}[scale=2]
      \inode[label=above:{$D = D_{t_0} = D_{u_0}$}]{a}{0, 0}
      \inode[label=left:$D_{t_1}$]{l1}{$(a) + (240:1)$}
      \inode[label=left:$D_{t_n}$]{l2}{$(l1) + (240:2)$}
      \inode[label=right:$D_{u_1}$]{r1}{$(a) + (300:1)$}
      \inode[label=right:$D_{u_m}$]{r2}{$(r1) + (300:2)$}
      \nnode{b}{$(l2) + (300:3)$}

      \draw[txn] (a) -- (l1) node[midway, left] {$t_1(D_{t_0})$};
      \draw[txn] (a) -- (r1) node[midway, right] {$u_1(D_{u_0})$};
      \draw[txn] (l1) -- (l2) node[midway, left] {
        $\begin{array}{l}
          t_2(D_{t_1}), \\
          \ldots, \\
          t_n(D_{t_{n-1}})
        \end{array}$
      };
      \draw[txn] (r1) -- (r2) node[midway, right] {
        $\begin{array}{l}
          u_2(D_{u_1}), \\
          \ldots, \\
          u_m(D_{u_{m-1}})
        \end{array}$
      };
      \draw[dtxn] (l2) -- (b) node[midway, left] {
        $\begin{array}{l}
          u_i(D_{u_{i-1}}), \\
          1 \leq i \leq m
        \end{array}$
      };
      \draw[dtxn] (r2) -- (b) node[midway, right] {
        $\begin{array}{l}
          t_i(D_{t_{i-1}}), \\
          1 \leq i \leq n
        \end{array}$
      };
    \end{tikzpicture}

@@

>>>>

A \emph{database state} is a total function $D: \var \to \ints$.  We say a
database state $D$ \emph{satisfies invariant} $I$, denoted $I(D)$, if $I$
evaluates to true after all variables in $I$ have been replaced by their
mapping in $D$. Let $\dbs$ denote the set of all database states.

@@

\subsection{\iconfluence{} Alternatives}

\newcommand{\ipreservation}{$\iinvariant$-preservation}
\newcommand{\ipreserving}{$\iinvariant$-preserving}

\newcommand{\isafety}{$\iinvariant$-safety}
\newcommand{\isafe}{$\iinvariant$-safe}

\newcommand{\istrength}{$\iinvariant$-strength}
\newcommand{\istrong}{$\iinvariant$-strong}

\newcommand{\istrengthstar}{\istrength$^\ast$}
\newcommand{\istrongstar}{\istrong$^\ast$}

\begin{itemize}
  \item \todo{\istrengthstar $\centernot \implies$ \iconfluence?}
  \item \todo{\iconfluence $\implies$ \istrengthstar?}
  \item \todo{Find something between \isafety{} and \iconfluence.}
  \item \todo{1-iconfluence with derived txns to start}
\end{itemize}

@@

  \item \textbf{k-\isafety.}
    $\forall D, D', B, C, c.\>
       C = \denote{B}(D) \land
       |B| \geq k \land
       I(B @ D') \implies
       I(C @ D) \land
       I(D \circ C \circ \denote{c}{D' \circ C})$

@@

  \item \textbf{1-\iconfluence.}
    $\forall D, c_1, c_2.\>
       I(D) \land
       I(D \circ \denote{c_1}(D)) \land
       I(D \circ \denote{c_2}(D)) \implies
       I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$

  \item \textbf{\istrengthstar{} \cite{gotsman2016cause}.}
    \leqnomode
    \begin{flalign*}
      & \exists G_0 \subseteq (\var \to \ints) \times (\var \to \ints).& \\
      & G_0(I) \subseteq I \land{} & \tag*{S2.} \\
      & \forall c, D, D'.\>
          D \in I \land
          (D, D') \in G_0^* \implies
          (D', \denote{c}(D)(D')) \in G_0 & \tag*{S3.}
    \end{flalign*}
    \reqnomode
    where $I$ is viewed as a set of database states.

  \item \textbf{\istrength{} \cite{gotsman2016cause}.}
    \leqnomode
    \begin{flalign*}
      & \exists G_0 \subseteq (\var \to \ints) \times (\var \to \ints). & \\
      & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
      & \forall c.\> & \tag*{T3.} \\
      & \exists P_1, \ldots, P_n \subseteq (\var \to \ints).\> & \\
      & \exists Q_1, \ldots, Q_n \subseteq (\var \to \ints).\> & \\
      & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
      & \quad \forall i \in \set{1, \ldots, n}.\> & \\
      & \quad \quad P_i \subseteq Q_i \land {} & \tag*{T3c.} \\
      & \quad \quad G_0(Q_i) \subseteq Q_i \land {} & \tag*{T3d.} \\
      & \quad \quad Q_i \times \denote{c}(P_i)(Q_i) \subseteq G_0 & \tag*{T3e.}
    \end{flalign*}
    \reqnomode
    where $I$ is viewed as a set of database states.
\end{itemize}

\begin{claim}
  0-\isafety $\implies$ \istrengthstar.
\end{claim}
\begin{proof}
  Consider an abitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
  respect to invariant $I$. We show $T$ is \istrongstar{} with respect to $I$.
  Let $G_0 = I \times I$. S2 holds trivially: $G_0(I) = I
  \subseteq I$. Consider an arbitrary $c$, $D$, and $D'$ where $D \in I$ and
  $(D, D') \in G_0^*$.
  \begin{align*}
    & D \in I \land (D, D') \in G_0^*  \\
    &\implies D \in I \land D' \in I \tag{Definition of $G_0$} \\
    &\implies D' \in I \land D' \circ \denote{c}(D) \in I \tag{0-\isafety} \\
    &\implies (D', D' \circ \denote{c}(D)) \in G_0 \tag{Definition of $G_0$}
  \end{align*}
\end{proof}

\begin{claim}
  \istrengthstar $\centernot\implies$ 0-\isafety
\end{claim}
\begin{proof}
  Consider a database over a single variable $x$. For convenience, we'll denote
  database states as integers. Let
  \begin{itemize}
    \item
      $G \defeq \set{(0, 0), (0, 2)} \cup \setst{(a, b)}{a \geq 2, b \geq a}$
    \item
      $I \defeq x = 0 \lor x \geq 2$, and
    \item
      the single \imp{} transaction $c \defeq
      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$.
  \end{itemize}

  First note that $c$ is \istrengthstar{} with respect to $I$. S2 holds
  trivially. To show S3 holds, consider an arbitrary pair $a, b \in G_0^*$.
  \begin{itemize}
    \item \textbf{Case $a = 0, b = 0$.}
      $0 \circ \denote{c}(0) = 0 + 0 = 0$ and $(0, 0) \in G_0$.
    \item \textbf{Case $a = 0, b >= 2$.}
      $b \circ \denote{c}(0) = b + 0 = b$ and $(b, b) \in G_0$.
    \item \textbf{Case $a \geq 2, b >= a$.}
      $b \circ \denote{c}(a) = b + a$ and $(b, b + a) \in G_0$.
  \end{itemize}

  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ and $\lnot I(1)$.
\end{proof}

@@

  \begin{tikzpicture}
    \node[] (0-isafety) at (0, 0) {0-\isafety};
    \node[] (1-isafety) at (0, -2) {1-\isafety};
    \node[] (k-isafety) at (0, -4) {$k$-\isafety};
    \node[] (istrengthstar) at (6, 0) {\istrengthstar};
    \node[] (ipreservation) at ($(0-isafety) + (30:3)$) {\ipreservation};
    \node[] (iconfluence) at ($(0-isafety) + (-30:3)$) {\iconfluence};
    \node[] (k-iconfluence) at ($(iconfluence) + (4, 0)$) {$k$-\iconfluence};
    \node[] (1-iconfluence) at ($(k-iconfluence) + (4, 0)$) {1-\iconfluence};

    \draw[impl] (0-isafety) -> (ipreservation);
    \draw[impl] (0-isafety) -> (iconfluence);
    \draw[impl] (iconfluence) -> (k-iconfluence);
    \draw[impl] (k-iconfluence) -> (1-iconfluence);
    \draw[impl] (0-isafety) -> (istrengthstar);
    \draw[impl] (istrengthstar) -> (iconfluence);
    \draw[impl] (istrengthstar) -> (ipreservation);
    \draw[impl] (0-isafety) -> (1-isafety);
    \draw[impl] (1-isafety) -> (k-isafety);
  \end{tikzpicture}

@@

  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{d1}{0, 0}
      \inode[label=above:$D'$]{d2}{1, 0}
      \dnode{b1}{0, -1}
      \nnode{b2}{1, -1}
      \draw[txn] (d2) -- (b2) node[midway, right]{$\denote{c}(D')$};
      \draw[dtxn] (d1) -- (b1) node[midway, left]{$\denote{c}(D')$};
    \end{tikzpicture}
    \caption{0-\isafety}
    \label{fig:0-isafety}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{a}{0, 0}
      \dnode{b}{0, -1}
      \draw[txn] (a) -- (b) node[midway, left]{$\denote{c}(D)$};
    \end{tikzpicture}
    \caption{\ipreservation}
    \label{fig:ipreservation}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{a}{0, 0}
      \inode[label=left:$D_{c}$]{l1}{$(a) + (240:1)$}
      \inode[label=right:$D_{d}$]{r1}{$(a) + (300:1)$}
      \dnode{b}{$(l1) + (300:1)$}
      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c}(D)$};
      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d}(D)$};
      \draw[txn] (l1) -- (b) node[midway, left] {$\denote{d}(D)$};
      \draw[txn] (r1) -- (b) node[midway, right] {$\denote{c}(D)$};
    \end{tikzpicture}
    \caption{1-\iconfluence}
    \label{fig:1-iconfluence}
  \end{subfigure}

@@

    \begin{tikzpicture}[scale=1.5]
      \inode[label=above:$D_0$]{l1}{0, 4}
      \dnode[label=left:$D_1$]{l2}{0, 3}
      \dnode[label=left:$D_2$]{l3}{0, 2}
      \dnode[label=left:$D_3$]{l4}{0, 1}
      \dnode{l5}{0, 0}

      \inode[label=above:$D_0'$]{r1}{1, 4}
      \inode[label=right:$D_1'$]{r2}{1, 3}
      \inode[label=right:$D_2'$]{r3}{1, 2}
      \inode[label=right:$D_3'$]{r4}{1, 1}
      \nnode{r5}{1, 0}

      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
      \draw[txn] (r2) -- (r3) node[midway, right]{$\denote{c_2}(D_1')$};
      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_3}(D_2')$};
      \draw[txn] (r4) -- (r5) node[midway, right]{$\denote{c_4}(D_3')$};

      \draw[txn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
      \draw[txn] (l2) -- (l3) node[midway, left]{$\denote{c_2}(D_1')$};
      \draw[txn] (l3) -- (l4) node[midway, left]{$\denote{c_3}(D_2')$};
      \draw[txn] (l4) -- (l5) node[midway, left]{$\denote{c_4}(D_3')$};
    \end{tikzpicture}

    \caption{$k$-\isafety, $k=3$}
    \label{fig:k-isafety}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering

    \begin{tikzpicture}[scale=1.5]
      \inode[label=above:$D$]{a}{0, 0}
      \inode[label=left:$D_{c_1}$]{l1}{$(a) + (240:1)$}
      \inode[label=left:$D_{c_2}$]{l2}{$(l1) + (240:1)$}
      \inode[label=left:$D_{c_3}$]{l3}{$(l2) + (240:1)$}

      \inode[label=right:$D_{d_1}$]{r1}{$(a) + (300:1)$}
      \inode[label=right:$D_{d_2}$]{r2}{$(r1) + (300:1)$}
      \inode[label=right:$D_{d_3}$]{r3}{$(r2) + (300:1)$}

      \dnode{b}{$(l3) + (300:3)$}

      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c_1}(D)$};
      \draw[txn] (l1) -- (l2) node[midway, left] {$\denote{c_2}(D_{c_1})$};
      \draw[txn] (l2) -- (l3) node[midway, left] {$\denote{c_3}(D_{c_2})$};
      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d_1}(D)$};
      \draw[txn] (r1) -- (r2) node[midway, right] {$\denote{d_2}(D_{d_1})$};
      \draw[txn] (r2) -- (r3) node[midway, right] {$\denote{d_3}(D_{d_2})$};
      \draw[txn] (l3) -- (b) node[midway, left] {$\denote{c_1, c_2, c_3}(D)$};
      \draw[txn] (r3) -- (b) node[midway, right] {$\denote{d_1, d_2, d_3}(D)$};
    \end{tikzpicture}

@@

  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{s2a}{0, 1}
      \dnode{s2b}{0, 0}
      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};

      \inode[label=above:$D$]{s3a}{2, 1}
      \nnode[]{s3b}{2, 0}
      \inode[label=above:$D'$]{s3c}{3, 1}
      \nnode[]{s3d}{3, 0}
      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
    \end{tikzpicture}
    \caption{\istrengthstar}
    \label{fig:istrengthstar}
  \end{subfigure}
